---
title: 'Peer-graded Assignment: Getting and Cleaning Data Course Project'
author: "JM""
date: "15 May 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Peer-graded Assignment: Getting and Cleaning Data Course Project

This is the codebook for the end of course assignment

#Read in the data and examine the structure

Following analysis of the zip file contents and the read-me documentation, the following files are read into R:

```{r}
#Read in all data
test <- read.table("test/X_test.txt")
train <- read.table("train/X_train.txt")
testlabels <- read.table("test/y_test.txt")
trainlabels <- read.table("train/y_train.txt")
activitylabels <- read.table("activity_labels.txt")
subjectlabelstrain <- read.table("train/subject_train.txt")
subjectlabelstest <- read.table("test/subject_test.txt")
features <- read.table("features.txt")

```

The structure of each data set is as follows:

```{r}
str(test)
str(train)
str(testlabels)
str(trainlabels)
str(activitylabels)
str(subjectlabelstest)
str(subjectlabelstrain)
str(features)
```

#Merging the Data Sets

From the above it can be seen that the test and training data sets ("test" and "train" need to be combined with the relevant activity and subject labels)

```{r pressure, echo=FALSE}
#Combine activity and subject labels with data sets
test <- cbind(testlabels,subjectlabelstest,test)
train <- cbind(trainlabels,subjectlabelstrain,train)
```

The train and test data sets have the same columns so they can be combined using the rbind function

```{r}
allData <- rbind(test,train)
```

#Re-labelling the Data Set Headers

The column headers are replaced with the variable names from the "features" object.

```{r}
colnames(allData) <- c("Activity","Subject",as.character(features[,2]))
```

#Filtering the Data Set Columns

Only measurements related to the mean or standard deviation are required. The grep function is used to extract these columns.

```{r}
meanSD <- cbind(allData[,1:2],allData[,grep("mean|std",colnames(allData))])
```

#Re-labelling the Data Set Activities

The activty names can be found in the "activitylabels" vector. This is merged with the main data set to replace the activity numbers

```{r}
meanSD <- merge(activitylabels,meanSD,by.x="V1",by.y="Activity",all.y=T)
meanSD <- meanSD[,-1]
names(meanSD)[1] <- "Activity"
```

#Creating Descriptive Variable Names

The strsplit, gsub, paste and substr functions are used to:
Split the header character strings into sub-parts, remove elements such as "()", format the axis value for each header, extract the time/frequency domain symbol (t or f) and replace with "time" or "frequency"

These elements are then recombined to give more meaningful header names

```{r}
getListElement <- function(x,n) {x[n]}
splitNames <- strsplit(colnames(meanSD)[3:ncol(meanSD)],"-")

activities <- sapply(splitNames,getListElement,n=1)

measures <- sapply(splitNames,getListElement,n=2)
measures <- gsub("\\()","",measures)

axis <- sapply(splitNames,getListElement,n=3)
axis[!is.na(axis)] <- paste(axis[!is.na(axis)],"Axis",sep="")
axis[is.na(axis)] <- ""

timeFreq <- substr(activities,1,1)
activityType <- substr(activities,2,nchar(activities))
timeFreq <- ifelse(timeFreq=="t","time","frequency")

names(meanSD)[3:ncol(meanSD)] <- 
  paste(measures,axis,activityType,timeFreq,sep="")
```

#Creating a Summary Table

The aggregate function is used to calculate the mean for each activity and subject. Finally the resulting table is ordered by subject and activity and exported as a .csv file.

```{r}
meanSD2<-aggregate(. ~Subject + Activity, meanSD, mean)
meanSD2<-meanSD2[order(meanSD2$Subject,meanSD2$Activity),]

write.csv(meanSD2,"TidyData.csv",row.names=F)
```
